A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-18 tex/body.tex
more

Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
For acyclicity, if $\ec{t}$ is a datatype equivalence class
whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
(as required by the precondition $x \in \FV( u )$),
then $\Ec$ is $\thD$-unsatisfiable.
For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
%for a capture-avoiding renaming of $\mu$-bound variables,
then $t \teq u$. This precondition may seem too restrictive, since
$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
value despite being $\alpha$-disequivalent, but the rule will
make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
\const{S}(\const{S}(y))$ must be equal to the entire term,
as demonstrated in the following example.

>>>>

Given the mapping $\Val$, the \rn{Acyclic} and \rn{Unique} rules operate as follows.
For acyclicity, if $\ec{t}$ is a datatype equivalence class
whose values $\Val \ec{t} = \MU x.\; u$ are cyclic
(as expressed by $x \in \FV( u )$),
then $\Ec$ is $\thD$-unsatisfiable.
For uniqueness, if $\ec{t}$, $\ec{u}$ are two codatatype equivalence classes
whose values $\Val \ec{t}$, $\Val \ec{u}$ are $\alpha$-equivalent,
%for a capture-avoiding renaming of $\mu$-bound variables,
then $t \teq u$. This requirement may seem too restrictive, since
$\MU x.\; \const{S}(x)$ and $\MU y.\; \const{S}(\const{S}(y))$ specify the same
value despite being $\alpha$-disequivalent, but the rule will
make some progress by discovering that the subterm $\const{S}(y)$ of $\MU y.\;
\const{S}(\const{S}(y))$ must be equal to the entire term,
as demonstrated next.
